Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-02-24 ThesisJP/Background/background.tex
Minar change in background.

\section {Medical ultrasound imaging}\label{sec:ultrasound}
Ultrasound imaging encompasses technology which generates images based on sound whose frequencies we can not hear. For medical ultrasound imaging frequencies between 2 MHz and 10 MHz are typically used. Frequency dependent attenuation force us to select a lower frequency if the object we want to image is located deep into the body. This is the reason why 2-4 MHz is used for cardiac imaging and a frequency around 7 MHz is used for vascular imaging. A piezoelectric ceramic is used to generate the ultrasonic frequencies, and is special in that way that ultrasound vibrations are generated when voltage is applied and \textit{vice versa}. It can therefore be used to both send and receive ultrasound signals, like a combined microphone and speaker. Piezoelectric elements organized in an array, is referred to as an ultrasound \textit{probe}, and a special probe is typically design for each medical imaging modality. For cardiac ultrasound imaging the probe footprint has to be smaller than the intercostal window in order to avoid shadowing by ribs. A line in an ultrasound image is generated by firing a ultrasound pulse into the body and sample the received echo. This amplitude trace is known as an A-mode image.

@@

If time delays are applied on the signal sent to the different elements it is possible to steer and focus the ultrasound beam in space. This technique is know as \textit{beamforming}, and for ultrasound imaging it is applied both when transmitting and receiving ultrasound pulses. For medical ultrasound imaging there are two main types of ultrasound probes, linear and phased arrays \cite{AngelUltrasound}. A cardiac ultrasound array is a phased array, which means that the image is constructed from beams, or A-mode images, which are gradually steered from $-\theta/2$ to $\theta/2$, resulting in sector which are $\theta$ wide. This kind of image is known as a B-mode image (B for brightness). If \nom{$M$}{Number of array elements} is the number of elements in the array, and \nom{$\Delta_m[n]$}{Per-element focusing and steering delays} is the per-element focusing and steering delay, the standard form of beamforming used on receive, known as \textit{delay-and-sum}, is given as:
\begin{align}\label{eq:das}
z[n] = \sum_{m = 0}^{M-1}w_m^*x_m[n - \Delta_m[n]] = \vec{w}^H\vec{x}[n],
\end{align}
where \nom{$\vec{w}$}{Array weight vector (apodization)} is know as the array weight vector or apodization, and $z[n]$ is the output A-mode image.  If $x_o$ is the phase center of our array and $p$ is the point where we want to steer and focus the ultrasound beam, we have to use the following delays:
\begin{align}
\Delta_m[n] =\frac{|x_m+p| - |x_o+p|}{c},
\end{align}
where \nom{$c$}{Speed of sound} is the speed of sound in human tissue. On transmit the inverse set of delays have to be used. Also note that on transmit we have to decide where to steer and focus prior to transmitting a pulse, transmit delays are therefore not time varying.

>>>>

\section {Medical ultrasound imaging}\label{sec:ultrasound}
Ultrasound imaging encompasses technology which generates images based on sound whose frequencies we can not hear. For medical ultrasound imaging frequencies between 2 MHz and 10 MHz are typically used, where frequency dependent attenuation force us to select a lower frequency if the object we want to image is located deep into the body. This is the reason why 2-4 MHz is used for cardiac imaging and a frequency around 7 MHz is used for vascular imaging. A piezoelectric ceramic is used to generate the ultrasonic frequencies, and is special in that way that ultrasound vibrations are generated when voltage is applied and \textit{vice versa}. It can therefore be used to both send and receive ultrasound signals, like a combined microphone and speaker. Piezoelectric elements organized in an array, is referred to as an ultrasound \textit{probe}, and a special probe is typically design for each medical imaging modality. For cardiac ultrasound imaging the probe must not only be designed for a certain frequency range, its footprint also needs to be smaller than the intercostal window in order to avoid shadowing by ribs. A line in an ultrasound image is generated by firing a ultrasound pulse into the body and sample the received echo. This amplitude trace is known as an A-mode image.

@@

If time delays are applied on the signal sent to the different elements it is possible to steer and focus the ultrasound beam in space. This technique is know as \textit{beamforming}, and for ultrasound imaging it is applied both when transmitting and receiving ultrasound pulses. For medical ultrasound imaging there are two main types of ultrasound probes, linear and phased arrays \cite{AngelUltrasound}. A cardiac ultrasound array is a phased array, which means that the image is constructed from beams, or A-mode images, which are gradually steered from $-\theta/2$ to $\theta/2$, resulting in sector which are $\theta$ wide. This kind of image is known as a B-mode image (B for brightness). If \nom{$M$}{Number of array elements} is the number of elements in the array, and \nom{$\Delta_m[n]$}{Per-element focusing and steering delays} is the per-element focusing and steering delay, the standard form of beamforming used on receive, known as \textit{delay-and-sum}, is given as:
\begin{align}\label{eq:das}
z[n] = \sum_{m = 0}^{M-1}w_m^*x_m[n - \Delta_m[n]] = \vec{w}^H\vec{x}[n],
\end{align}
where \nom{$\vec{w}$}{Array weight vector (apodization)} is know as the array weight vector or apodization, and $z[n]$ is the output A-mode image.  If $p_o$ is the phase center of our array and $p$ is the point where we want to steer and focus the ultrasound beam, we have to use the following delays:
\begin{align}
\Delta_m[n] =\frac{|p_m+p| - |p_o+p|}{c},
\end{align}
where \nom{$c$}{Speed of sound} is the speed of sound in human tissue, and $p_m$ is the position of the m'th element. On transmit the inverse set of delays have to be used. Also note that on transmit we have to decide where to steer and focus prior to transmitting a pulse, transmit delays are therefore not time varying.
